---
title: "Biodiversity Module"
author: <a href="http://bigdata.duke.edu/data">Duke Data+</a> Interactive Environmental Data Team
date: ""
output: 
  html_document:
    theme: journal
---

```{r, echo=FALSE}
mythical<- read.csv("data/mythical.csv")
m<- as.character(mythical[,1])
getCommunity <- function(S=10,N=1000,equalprob=T){
  if(S > 51) S <- 51 # right now we only have 26 species - each unique letter in the alphabet
  prob <- rep(1/S,S)
  if(!equalprob){
    Nx <- rpois(S,sample(1:S,1))
    prob <- Nx/sum(Nx)    
  }
  sub<-m[sample(S,N,replace=T,prob=prob)]
  print(sub)
  return (sub)
}
```

## Background

Biodiversity indices are used by environmental scientists to measure the diversity of species in a community.

Three common measures are:

* Richness
* Shannon index
* Simpson index

We can calculate these indices with a data set of observations. For example, here is a hypothetical community (for new communities, just reload this page):

```{r, echo=FALSE}
dat <- sample(m,20,replace=T)
dat
```

## Richness

Richness ($S$) is the number of unique species. With a list of organism observations (above), the richness is:

```{r, echo=FALSE}
S <- length(unique(dat))
paste(S, "unique species")
```


## Shannon index

The Shannon index measures the evenness of species observations with relative proportions. The index is:

$H'=- \sum_{i=1}^S p_i \ln p_i$

where $S$ is the number of unique species and $p_i$ is the proportion of the $i^{th}$ species, which is $\frac{n_i}{N}$, where $N$ is the total number of individuals.


```{r, echo=FALSE}
Pi <- table(dat)/length(dat)

H <- -sum(Pi*log(Pi))
H
```

Higher values have higher evenness, and the maximum value is the log of the species richness ($\ln S$).   For this example: the maximum value is:

```{r, echo=FALSE}
X<- log(S)
X
```

Therefore, this population has a high biodiversity

## Simpson index

The Simpson index measures the concentration of observations in certain groups. This index is:

$\lambda=\sum_{i=1}^S p_i^2$

It measures the probability that two individuals randomly chosen from a sample will have the same species.

For our example data set:

```{r, echo=FALSE}
lambda <- sum(Pi^2)
paste("Simpson is:", lambda)
```

It ranges between 0 and 1, and lower numbers are more diverse. Because of this inverse ranking, some people refer to the inverse of the Simpson index:

$D = 1/\lambda$

```{r, echo=FALSE}
D <- 1/lambda
paste("Inverse Simpson is:",  D)
```

Here, 1 is the lowest number and higher values have higher diversity.


